---
title: "Industrial sewage network discharge"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F, message=F, results=F, dev='svg')

library(tidyverse) #loads multiple packages (see https://tidyverse.tidyverse.org/)

#core tidyverse packages loaded:
# ggplot2, for data visualisation. https://ggplot2.tidyverse.org/
# dplyr, for data manipulation. https://dplyr.tidyverse.org/
# tidyr, for data tidying. https://tidyr.tidyverse.org/
# readr, for data import. https://readr.tidyverse.org/
# purrr, for functional programming. https://purrr.tidyverse.org/
# tibble, for tibbles, a modern re-imagining of data frames. https://tibble.tidyverse.org/
# stringr, for strings. https://stringr.tidyverse.org/
# forcats, for factors. https://forcats.tidyverse.org/
# lubridate, for date/times. https://lubridate.tidyverse.org/

#also loads the following packages (less frequently used):
# Working with specific types of vectors:
#     hms, for times. https://hms.tidyverse.org/
# Importing other types of data:
#     feather, for sharing with Python and other languages. https://github.com/wesm/feather
#     haven, for SPSS, SAS and Stata files. https://haven.tidyverse.org/
#     httr, for web apis. https://httr.r-lib.org/
#     jsonlite for JSON. https://arxiv.org/abs/1403.2805
#     readxl, for .xls and .xlsx files. https://readxl.tidyverse.org/
#     rvest, for web scraping. https://rvest.tidyverse.org/
#     xml2, for XML. https://xml2.r-lib.org/
# Modelling
#     modelr, for modelling within a pipeline. https://modelr.tidyverse.org/
#     broom, for turning models into tidy data. https://broom.tidymodels.org/

# Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors

#loading additional relevant packages
library(cowplot) #for plot_grid() and get_legend

#path for data 
path_source <- "source_data/0_industry_discharge/georisque_data/"

#setting graphs theme
theme_set(
  theme_minimal() +
    theme(
      text = element_text(family = "Times New Roman"),
      plot.title = element_text(face="bold")
      )
  )

#setting viridis theme for colors
scale_colour_continuous <- scale_colour_viridis_c
scale_colour_discrete   <- scale_colour_viridis_d
scale_colour_binned     <- scale_colour_viridis_b
#setting viridis theme for fill
scale_fill_continuous <- scale_fill_viridis_c
scale_fill_discrete   <- scale_fill_viridis_d
scale_fill_binned     <- scale_fill_viridis_b

Source <- "Source: géorisque database\ncomputation by Thomas Starck"

# Load the function file
source("functions.R")
```

# Source, description and loading data

The georisque database is public and reports industrial discharge, when they are made compulsory by the reglementary thresholds. The GEREP database is not open access but was provided by the French administration ; it contains more industry discharges than simply the mandatory reported ones.

::: panel-tabset

## Load géorisque data

::: panel-tabset

### Source and description

The data were downloaded on the [georisque](https://www.georisques.gouv.fr/installations-industrielles-rejetant-des-polluants) website, reporting industrial pollution discharges. Read the description of the data on [this page](https://www.georisques.gouv.fr/risques/registre-des-emissions-polluantes).

All pollution flows are reported in kg/year in the database.

### Emissions

We load the emission file emissions.csv, for each year. They report the facility ID and name, pollutant type (145 different), discharge environment (air, water direct, water indirect, ground), and quantity in kg/an. We focus only on NGL, Pt, DBO5, DCO, MES.

```{r loading_emissions}
#loading pollutant data
file_emission <- list.files( #read and merge csv of all years
    path = path_source,
    pattern = "emissions.csv", full.names = T, recursive = T) %>%
    lapply(read_csv2) %>% bind_rows()

# see the 145 different pollutants
unique(file_emission$polluant)

# focus only on NGL, Pt, DBO5, DCO, MES
file_emission <- file_emission %>% 
  filter(
    polluant %in% c(
      "Phosphore total",
      "Azote total",
      "Demande chimique en oxygène (DCO)",
      "Demande biologique en oxygène (DBO5)",
      "Matières en suspension (MES)"
      )
    ) %>%
  mutate(
    quantite = as.numeric(quantite)
  )
```

We load the files describing the facilities (etablissements.csv). They report the site name, address, coordinates, the European Pollutant Release and Transfer Register code (EPRTR code and wording), and the APE code and wording. EPRTR categories are listed [here](https://www.georisques.gouv.fr/risques/registre-des-emissions-polluantes/synthese/donnees#/recherche=secteur), and APE codes are listed [here](https://www.georisques.gouv.fr/risques/registre-des-emissions-polluantes/etablissement/donnees#/secteur=codeAPE)

We merge this emissions file through their unique ID. There are 200 APE codes and 50 EPRTR codes related to our pollution of interest (NGL, Pt, DBO5, DCO, MES).

```{r loading_sites_ID}
year_min <- 2003
year_max <- 2020
f_load_id_sites <- function(year_min, year_max){
  temp <- 
    read_csv2(paste0(path_source, year_min, "/etablissements.csv")) %>% 
    mutate(
      coordonnees_x = as.numeric(coordonnees_x),
      coordonnees_y = as.numeric(coordonnees_y),
      annee_emission = year_min
      )
  for (i in (year_min+1):year_max){
    temp2 <- 
      read_csv2(paste0(path_source, i, "/etablissements.csv")) %>% 
      mutate(
        coordonnees_x = as.numeric(coordonnees_x),
        coordonnees_y = as.numeric(coordonnees_y),
        annee_emission = i
        )
    temp <- bind_rows(temp, temp2)
  }
  return(temp)
}

file_id_sites <- f_load_id_sites(2003, 2020)

file_emissions_id_site <- left_join(
  file_emission, file_id_sites, by=c("annee_emission","identifiant")
)

# see the 300 APE codes
unique(file_emissions_id_site$libelle_ape)

# see the 60 EPRTR codes
unique(file_emissions_id_site$libelle_eprtr)

# we create the main file
georisque <- file_emissions_id_site %>%
  spread(polluant, quantite)

#we can also get commune, dep and region ID
georisque <- georisque %>%
  select(
    Year = annee_emission,
    Id = identifiant,
    NGL = `Azote total`,
    Pt = `Phosphore total`,
    DCO = `Demande chimique en oxygène (DCO)`,
    DBO5 = `Demande biologique en oxygène (DBO5)`,
    MES = `Matières en suspension (MES)`,
    site_name = nom_etablissement.x,
    SIRET_code = numero_siret,
    discharge_environment = milieu,
    lat_site = coordonnees_x,
    long_site = coordonnees_y,
    code_epsg,
    code_ape,
    libelle_ape,
    code_eprtr,
    libelle_eprtr
  )

# we change the french wording of discharge environment to English
georisque <- georisque %>%
  mutate(
    discharge_environment = case_when(
      discharge_environment == "Eau (direct)"~ "water (direct)",
      discharge_environment == "Eau (indirect)"~ "water (indirect)",
      discharge_environment == "Sol"~ "ground",
    )
  )

#see a summary of emissions according to the environmental discharge
table(
  file_emissions_id_site %>% 
    filter(
      polluant %in% c(
      "Phosphore total",
      "Azote total",
      "Demande chimique en oxygène (DCO)",
      "Demande biologique en oxygène (DBO5)",
      "Matières en suspension (MES)"
      )
    ) %>%
    select(polluant, milieu)
  )
```

### Discharged flows

For 2019 and 2020, for each site (ID and name), there are files (rejets.csv) specifying if there are some discharges in the network or in the environment, in m3/y.

```{r}
file_discharge <- list.files( #read and merge csv of all years
    path = path_source,
    pattern = "rejets.csv", full.names = T, recursive = T) %>%
    lapply(
      read_csv2
      ) %>% 
  bind_rows()

file_discharge <- file_discharge %>%
  rename(
    Year = annee_rejet,
    Id = identifiant,
    site_name = nom_etablissement,
    discharge_network_m3_y = rejet_raccorde_m3_par_an,
    discharge_environment_m3_y = rejet_isole_m3_par_an
    
  )
```

:::

## Load GEREP data

We load the data.

```{r}
path_source <- "source_data/0_industry_discharge/GEREP_data/"

file_GEREP <- readxl::read_excel(paste0(path_source, "2010_2021_Rejets raccordés eau par agence.xlsx"))

GEREP <- file_GEREP %>%
  select(
    Year = Année,
    basin = `Agence de l'eau`,
    nutrient = Substance,
    incoming = `Masse émise retenue (kg/an)`,
    discharged = `Rejet final après épuration (kg/an)`,
    INSEE_REG = `Code INSEE Région`,
    name_REG = Région,
    INSEE_DEP = `Code INSEE département`,
    name_DEP = Département,
    INSEE_COM = `Code INSEE Commune`,
    name_COM = Commune
  )

# Simplify name of the basin
GEREP <- GEREP %>%
  mutate(
    basin = case_when(
      basin == "Rhône-Méditerranée-Corse" ~ "Rhône-Méditerranée",
      T ~ basin
    )
  )
```

We compute the flows at the national scale

```{r}
temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  group_by(Year, nutrient) %>%
  summarise(
    incoming = signif(sum(incoming, na.rm=T)/10^6, 3),
    discharged = signif(sum(discharged, na.rm=T)/10^6, 3),
    nb_facilities = sum(is.na(nutrient)==F)
    )

temp_in <- temp %>% select(-c(discharged, nb_facilities)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming, nb_facilities)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )

temp_nb <- temp %>% select(-c(incoming, discharged)) %>%
  spread(nutrient, nb_facilities) %>%
  rename(
    NGL_nb_facilities = `Azote total (N)`,
    Pt_nb_facilities = `Phosphore total (P)`,
    DBO5_nb_facilities = `Demande biologique en oxygène (DBO5)`,
    DCO_nb_facilities = `Demande chimique en oxygène (DCO)`,
    MES_nb_facilities = `Matières en suspension (MES)`
  )

GEREP_national <- 
  left_join(
    temp_in, temp_out, by="Year"
)
GEREP_national <-   left_join(
    GEREP_national, temp_nb, by="Year"
)
```

We compute the flows for each basin

```{r}
temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  group_by(Year, basin, nutrient) %>%
  summarise(
    incoming = signif(sum(incoming, na.rm=T)/10^6, 3),
    discharged = signif(sum(discharged, na.rm=T)/10^6, 3),
    nb_facilities = sum(is.na(nutrient)==F)
    )

temp_in <- temp %>% select(-c(discharged, nb_facilities)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming, nb_facilities)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )

temp_nb <- temp %>% select(-c(incoming, discharged)) %>%
  spread(nutrient, nb_facilities) %>%
  rename(
    NGL_nb_facilities = `Azote total (N)`,
    Pt_nb_facilities = `Phosphore total (P)`,
    DBO5_nb_facilities = `Demande biologique en oxygène (DBO5)`,
    DCO_nb_facilities = `Demande chimique en oxygène (DCO)`,
    MES_nb_facilities = `Matières en suspension (MES)`
  )

GEREP_basins <- left_join(
  temp_in, temp_out, by=c("Year", "basin")
)
GEREP_basins <- left_join(
  GEREP_basins, temp_nb, by=c("Year", "basin")
)
```

We aslo create a the file with the flows of all the facilities

```{r}
temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  mutate(ID = 1:n())

temp_in <- temp %>% select(-c(discharged)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )


GEREP_all <- left_join(
  temp_in, 
  temp_out %>% select(ID, NGL_out, Pt_out, DBO5_out, DCO_out, MES_out), by="ID"
)
```

## More information

About [INERIS](https://www.ineris.fr/fr), responsible for IREP (installation registre émissions polluantes).

About [Aida](https://aida.ineris.fr/recherche?search=IREP), INERIS website on risk prevention and environmental protection regulations, and the related [decree](https://aida.ineris.fr/reglementation/arrete-261212-modifiant-larrete-31-janvier-2008-relatif-registre-a-declaration).

About the European registry [E-PRTR](https://ec.europa.eu/environment/industry/stationary/eper/legislation.htm).

:::

# géorisque data 

On [this page](https://www.georisques.gouv.fr/risques/registre-des-emissions-polluantes/synthese/donnees#/recherche=secteur) we can see that what is called water indirect ("eau indirect") in the data base corresponds to pollutants sent to wastewater treatment plants ("eau vers station d'épuration"). However it is unclear whether the reported pollutant flows correspond to the quantity discharged to the sewage network or the quantity discharge to the environment after WWTP end of pipe pollution treatment.

::: panel-tabset

## Outliers identification {.tabset .tabset-fade}

Spikes in the right chart indicate some potential incoherent reported pollution quantities, that we investigate.

```{r}
f_graph <- function(nutrient){
  temp <- georisque %>%
    filter(is.na(!!as.symbol(nutrient))==F) %>%
    group_by(Year) %>%
    summarise(
      n=n(),
      nutrient = sum(!!as.symbol(nutrient))/10^6, #kt
    )
  temp2 <- georisque %>%
    filter(is.na(!!as.symbol(nutrient))==F) %>%
    group_by(Year, discharge_environment) %>%
    summarise(
      n=n(),
      nutrient = sum(!!as.symbol(nutrient))/10^6, #kt
    )
  
  g1 <- ggplot(temp2) + 
      geom_area(aes(Year, n, fill=discharge_environment), alpha=.8) + 
      geom_line(data = temp, aes(Year, n)) +
      ylim(0, NA) +
      labs(
        title = paste(nutrient, "industrial emissions"),
        subtitle = paste("nb of industrial sites reporting", nutrient, "emissions"),
        x="", y="",
        caption = ""
      ) +
      theme(
        legend.position = "none"
      )
  g2 <- ggplot(temp2) + 
      geom_area(aes(Year, nutrient, fill=discharge_environment), alpha=.8) + 
      geom_line(data = temp, aes(Year, nutrient)) +
      ylim(0, NA) +
      labs(
        title = "",
        subtitle = paste("total", nutrient, "industrial emissions (kt/year)"),
        x="", y="",
        caption = Source,
        fill="environmental discharge:"
      ) +
      theme(
        legend.position = "bottom"
      )
  leg <- get_legend((g2))
  
  plot_grid(
    plot_grid(g1, g2 + theme(legend.position = "none"), align = "hv"),
    leg,
    nrow=2, rel_heights = c(0.9, 0.1)
  )
}
```

::: panel-tabset

### NGL

Outliers : in 2017 for indirect water discharge ; in 2010 for direct water discharge (and in 2007 ?).

```{r}
f_graph("NGL")
```

### Pt

Outliers : in 2010 direct water discharge ; in 2008 and 2020 ground discharge.

```{r}
f_graph("Pt")
```

### DBO5

Outliers : in 2010 direct water discharge, and in 2007.

```{r}
f_graph("DBO5")
```

### DCO

Outliers : in 2010 direct water discharge ; in 2018, 2019 and 2020 indirect water discharge.

```{r}
f_graph("DCO")
```

### MES

Outliers : in 2010 direct water discharge ; in 2018, 2019 and 2020 indirect water discharge.

```{r}
f_graph("MES")
```

### Outliers correction

For the 5 nutrients, in 2010 for SIAAP Marne Aval (ID 0079305101)

-   NGL, Pt factor 100
-   DCO, DBO5 and MES factor 1000

```{r}
#NGL
georisque$NGL[georisque$Id == "0079305101" & georisque$Year == 2010] <- 
  georisque$NGL[georisque$Id == "0079305101" & georisque$Year == 2010]/100
#Pt
georisque$Pt[georisque$Id == "0079305101" & georisque$Year == 2010] <- 
  georisque$Pt[georisque$Id == "0079305101" & georisque$Year == 2010]/100
#DBO5
georisque$DBO5[georisque$Id == "0079305101" & georisque$Year == 2010] <- 
  georisque$DBO5[georisque$Id == "0079305101" & georisque$Year == 2010]/1000
#DCO
georisque$DCO[georisque$Id == "0079305101" & georisque$Year == 2010] <- 
  georisque$DCO[georisque$Id == "0079305101" & georisque$Year == 2010]/1000
#MES
georisque$MES[georisque$Id == "0079305101" & georisque$Year == 2010] <- 
  georisque$MES[georisque$Id == "0079305101" & georisque$Year == 2010]/1000
```

NGL :

-   2007, Id 0006506939, SIAAP : OK, juste change names (SIAAP - Seine Aval) and ID starting 2008, no need to correct
-   2017, Id 0005403171, SMET NORD EST 71 : only year reported =\> we remove it

```{r}
georisque$NGL[georisque$Id == "0005403171" & georisque$Year == 2017] <- NA
```

Pt :

-   2008, Id 0005103644 SOVALD : only year reported =\> we remove it
-   2020, Id 0005106528 SUEZ ORGANIQUE : only year reported =\> we remove it

```{r}
georisque$Pt[georisque$Id == "0005103644" & georisque$Year == 2008] <- NA
georisque$Pt[georisque$Id == "0005106528" & georisque$Year == 2020] <- NA
```

DBO5 :

-   2007, Id 0009000031, Base Aérienne 103 CDT René MOUCHOTTE : only year reported =\> we remove it

```{r}
georisque$DBO5[georisque$Id == "0009000031" & georisque$Year == 2007] <- NA
```

DCO :

-   2018, 2019 and 2020, Id 0006402259 SERAMM, Usine des boues : about 10 times higher than the previous years (same with MES but not with DBO5 which is coherent). We divide the values of these years by 10.

```{r}
georisque$DCO[georisque$Id == "0006402259" & georisque$Year %in% c(2018, 2019, 2020)] <- 
  georisque$DCO[georisque$Id == "0006402259" & georisque$Year %in% c(2018, 2019, 2020)]/10
```

MES :

-   2019, Id 0006700018, Matthieu COTTING : only year reported =\> we remove it
-   2018, 2019, 2020, Id 0006402259 SERAMM Usine des boues : about 10 times higher than the previous years (same with DCO but not with DBO5, which is coherent). We divide by 10.

```{r}
georisque$MES[georisque$Id == "0006700018" & georisque$Year ==2019] <- NA

georisque$MES[georisque$Id == "0006402259" & georisque$Year %in% c(2018, 2019, 2020)] <- 
  georisque$MES[georisque$Id == "0006402259" & georisque$Year %in% c(2018, 2019, 2020)]/10
```

:::

## Corrected Flows {.tabset .tabset-fade}

Here are the pollutant flows after outliers correction. For the 5 nutrients, the data seem reliable starting 2008.

::: panel-tabset

### NGL

```{r}
f_graph("NGL")
```

### Pt

```{r}
f_graph("Pt")
```

### DBO5

```{r}
f_graph("DBO5")
```

### DCO

```{r}
f_graph("DCO")
```

### MES

```{r}
f_graph("MES")
```

:::

## Understand network discharge {.tabset .tabset-fade}

For 2019 and 2020 discharge flows (in m3/year) in the sanitation network and in the environment are reported in the emission.csv file. We combine these data with the one presented in the previous Emission flows part, to check for coherence. Normally the nutrient flows discharged in the network should correspond to the "water (indirect)" flows of the Emission flows part.

::: panel-tabset

```{r}
merge <- inner_join(
  file_discharge %>% 
    #We only keep non null discharge flows into networks
    filter(is.na(discharge_network_m3_y)==F) %>%
    filter(discharge_network_m3_y != 0),
  georisque, 
  by = c("Year", "Id")
)

# check the in the 2 databases the site names are the same
nrow(merge %>% filter(site_name.x != site_name.y))
# 0 lines, OK, so we only keep one site_name
merge <- merge %>% 
  select(-site_name.y) %>%
  rename(site_name = site_name.x)
```

Some of what is reported as "connected disharged" in the emission file is actually not a discharge into the sanitation network, but a discharge in the environment from waste water treatment plant. We identify the waste water treatment plants by their APE code ("Collecte et traitement des eaux usées") and EPRTR code ("Installations de traitement des eaux urbaines résiduaires d'une capacité de 100 000 équivalents habitants").

```{r}
unique(merge$libelle_ape)
unique(merge$libelle_eprtr)

merge <- merge %>% 
  mutate(
    WWTP = case_when(
      libelle_ape == "Collecte et traitement des eaux usées" | 
        libelle_eprtr == "Installations de traitement des eaux urbaines résiduaires d'une capacité de 100 000 équivalents habitants" ~
        "WWTP", 
      T ~ "not WWTP"
    )
  )
```

In the graphs below, we can see that besides what looks like some little incoherences, the "indirect water discharge" (black line) corresponds well to the discharge into networks of facilities that are not WWTP (columns for 2019 and 2020, left).

So in the following, we will use the "indirect water discharge" as a proxy for what is discharged into the sanitation networks.

```{r}
f_graph_comparison <- function(nutrient_select){
  #in the main file keep only indirect water discharge of NGL, normally corresponding to discharge in sanitation networks.
  #agregate at the nationa scale
  temp <- georisque %>%
    filter(
      is.na(!!as.symbol(nutrient_select))==F & discharge_environment == "water (indirect)"
    ) %>%
    group_by(Year) %>%
    summarise(nutrient=sum(!!as.symbol(nutrient_select)))
  
  #keep only non null NGL emissions
  temp2 <- merge %>% 
    filter(is.na(!!as.symbol(nutrient_select))==F) %>%
    rename(nutrient = !!as.symbol(nutrient_select))
  
  ggplot(temp2) +
    geom_col(
      aes(
        Year, nutrient/10^6, 
        fill= discharge_environment
        ),
      alpha=.8
      ) +
    geom_line(
      data=temp %>% mutate(WWTP = "not WWTP"),
      aes(Year, nutrient/10^6)
      ) +
    facet_wrap(vars(WWTP)) +
    labs(
      x="", y="kt/year",
      title = paste("Comparison of", nutrient_select, "'network discharge' to 'indirect water discharge'"),
      subtitle = "network discharge of WWTP (right) and other facilities (left)\nwater discharged in sewage networks (columns) and indirect water discharge (line)",
      caption=Source,
      fill = "network discharge to"
    )
  
}
```

### NGL

```{r}
f_graph_comparison("NGL")
```

### Pt

```{r}
f_graph_comparison("Pt")
```

### DCO

```{r}
f_graph_comparison("DCO")
```

### DBO5

```{r}
f_graph_comparison("DBO5")
```

### MES

```{r}
f_graph_comparison("MES")
```

:::

## Final data {.tabset .tabset-fade}

We save the sewage network discharge for all the undustrial sites.

```{r}
temp <- georisque %>%
  filter(
    discharge_environment=="water (indirect)"
    )

output_data <- "output_data/industry_network_discharge/"
f_save_csv_files(
  temp, 
  output_data, 
  "industry_network_discharge_georisque_ALL.csv"
  )
```

We also save the data agregated at the national scale. Below are the grahps for each nutrient.

```{r}
national_network_discharge <- georisque %>%
  filter(
    discharge_environment=="water (indirect)"
    ) %>%
  group_by(Year) %>%
  summarise(
    across(
      c(Pt, NGL, DBO5, DCO, MES), 
      ~ signif(sum(.x, na.rm=T)/10^6, 3) #converted in kt per year
    )
  )

temp <- georisque %>%
  filter(
    discharge_environment=="water (indirect)"
    ) %>%
  group_by(Year) %>%
  summarise(
    across(
      c(Pt, NGL, DBO5, DCO, MES), 
      ~ sum(is.na(.x)==F), .names = "{.col}_nb_facilities" #converted in kt per year
    )
  )

national_network_discharge <- 
  left_join(national_network_discharge, temp, by="Year")

output_data <- "output_data/industry_network_discharge/"
f_save_csv_files(
  national_network_discharge, 
  output_data, 
  "industry_network_discharge_georisque_national.csv"
  )
```

```{r}
f_graph_final <- function(dataset, nutrient_flow, nutrient_nb_facilities){
  g <- plot_grid(
    ggplot(dataset) +
      geom_area(aes(Year, !!as.symbol(nutrient_flow), fill=""), alpha=.8) +
        theme(legend.position = "none") +
        labs(
          x="", y=paste("kt of", nutrient_flow, "per year"), 
          caption = "\n",
          title="Industrial discharge to sewage network in France",
          subtitle = paste(nutrient_flow, "flow")
        ),
      ggplot(dataset) +
        geom_area(aes(Year, !!as.symbol(nutrient_nb_facilities), fill=""), alpha=.8) +
        theme(legend.position = "none") +
        labs(
          x="", y="", 
          caption = Source,
          title="",
          subtitle = paste("nb of facilities reporting",  nutrient_flow, "discharge")
        )
    )
  return(g)
}

```

::: panel-tabset

### Pt

Order of magnitude for the 2010s : 0.7 ktP/year.

```{r}
f_graph_final(national_network_discharge, "Pt", "Pt_nb_facilities")
```

### NGL

Order of magnitude for the 2010s : 3-4 ktN/year.

```{r}
f_graph_final(national_network_discharge, "NGL", "NGL_nb_facilities")
```

### DBO5

Order of magnitude for the 2010s : 60-70 ktDBO5/year.

```{r}
f_graph_final(national_network_discharge, "DBO5", "DBO5_nb_facilities")
```

### DCO

Order of magnitude for the 2010s : 150 ktDCO/year.

```{r}
f_graph_final(national_network_discharge, "DCO", "DCO_nb_facilities")
```

### MES

Order of magnitude for the 2010s : 10-15 ktMES/year.

```{r}
f_graph_final(national_network_discharge, "MES", "MES_nb_facilities")
```

:::

:::

# GEREP data 

::: panel-tabset

## Before Correction {.tabset .tabset-fade}

<details>

<summary>see outliers</summary>

PHOSPHORUS

-   in 2017 for CHAGNY, in and out, factor 100 (incoming==55879, there are several Chagny)
-   in 2010 and 2011 for PORT-SUR-SAONE, does not exist afterwards. We did not correct it for now.
-   in 2010 and 2011 for VITRY-SUR-SEINE, factor 100 in, factor 10 out

NGL

-   in 2017 for CHAGNY, factor 100 (incoming == 7711302, there are several Chagny)
-   in 2019 for AUBIGNY, factor 1000 in, factor 10 000 out
-   in 2013 for ILLZACH, factor 10 in and out

DCO

-   in 2018 for MARSEILLE, factor 10 in and out 10 (incoming==94603000, there are several Marseille)

MES

-   in 2018, MARSEILLE, factor 10 in and out 10 (incoming==24927000, there are several Marseille)
-   in 2019 for BISCHWILLER, factor 10 000 in and out

</details>

```{r}
#PHOSPHORUS
#CHAGNY in, 2017
GEREP$incoming[
  GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Phosphore total (P)" & GEREP$incoming==55879
  ] <- 
  GEREP$incoming[
    GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Phosphore total (P)" & GEREP$incoming==55879
    ]/100
#CHAGNY out, 2017
GEREP$discharged[
  GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Phosphore total (P)" & GEREP$incoming==55879
  ] <- 
  GEREP$discharged[
    GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Phosphore total (P)" & GEREP$incoming==55879
    ]/100
#VITRY-SUR-SEIN in, 2010 and 2011
GEREP$incoming[
  GEREP$Year%in%c(2010, 2011) & GEREP$name_COM == "VITRY-SUR-SEINE" & GEREP$nutrient=="Phosphore total (P)"
  ] <- 
  GEREP$incoming[
    GEREP$Year%in%c(2010, 2011) & GEREP$name_COM == "VITRY-SUR-SEINE" & GEREP$nutrient=="Phosphore total (P)"
    ]/100
#VITRY-SUR-SEIN out, 2010 and 2011
GEREP$discharged[
  GEREP$Year%in%c(2010, 2011) & GEREP$name_COM == "VITRY-SUR-SEINE" & GEREP$nutrient=="Phosphore total (P)"
  ] <- 
  GEREP$discharged[
    GEREP$Year%in%c(2010, 2011) & GEREP$name_COM == "VITRY-SUR-SEINE" & GEREP$nutrient=="Phosphore total (P)"
    ]/10


#NITROGEN
#CHAGNY in, 2017
GEREP$incoming[
  GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Azote total (N)" & GEREP$incoming==7711302
  ] <- 
  GEREP$incoming[
    GEREP$Year==2017 & GEREP$name_COM == "CHAGNY" & GEREP$nutrient=="Azote total (N)" & GEREP$incoming==7711302
    ]/100
#AUBIGNY in, 2019
GEREP$incoming[
  GEREP$Year==2019 & GEREP$name_COM == "AUBIGNY" & GEREP$nutrient=="Azote total (N)"
  ] <- 
  GEREP$incoming[
    GEREP$Year==2019 & GEREP$name_COM == "AUBIGNY" & GEREP$nutrient=="Azote total (N)"
    ]/1000
#AUBIGNY out, 2019
GEREP$discharged[
  GEREP$Year==2019 & GEREP$name_COM == "AUBIGNY" & GEREP$nutrient=="Azote total (N)"
  ] <- 
  GEREP$discharged[
    GEREP$Year==2019 & GEREP$name_COM == "AUBIGNY" & GEREP$nutrient=="Azote total (N)"
    ]/10000
#ILLZACH in, 2013
GEREP$incoming[
  GEREP$Year==2013 & GEREP$name_COM == "ILLZACH" & GEREP$nutrient=="Azote total (N)"
  ] <- 
  GEREP$incoming[
    GEREP$Year==2013 & GEREP$name_COM == "ILLZACH" & GEREP$nutrient=="Azote total (N)"
    ]/10
#ILLZACH out, 2013
GEREP$discharged[
  GEREP$Year==2013 & GEREP$name_COM == "ILLZACH" & GEREP$nutrient=="Azote total (N)"
  ] <- 
  GEREP$discharged[
    GEREP$Year==2013 & GEREP$name_COM == "ILLZACH" & GEREP$nutrient=="Azote total (N)"
    ]/10


#DCO
#MARSEILLE in, 2018 
GEREP$incoming[
  GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Demande chimique en oxygène (DCO)" & GEREP$incoming==94603000
  ] <- 
  GEREP$incoming[
    GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Demande chimique en oxygène (DCO)" & GEREP$incoming==94603000
    ]/10
#MARSEILLE out, 2018 
GEREP$discharged[
  GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Demande chimique en oxygène (DCO)" & GEREP$incoming==94603000
  ] <- 
  GEREP$discharged[
    GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Demande chimique en oxygène (DCO)" & GEREP$incoming==94603000
    ]/10

#MES
#MARSEILLE in, 2018 
GEREP$incoming[
  GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Matières en suspension (MES)" & GEREP$incoming==24927000
  ] <- 
  GEREP$incoming[
    GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Matières en suspension (MES)" & GEREP$incoming==24927000
    ]/10
#MARSEILLE out, 2018 
GEREP$discharged[
  GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Matières en suspension (MES)" & GEREP$incoming==24927000
  ] <- 
  GEREP$discharged[
    GEREP$Year==2018 & GEREP$name_COM == "MARSEILLE" & GEREP$nutrient=="Matières en suspension (MES)" & GEREP$incoming==24927000
    ]/10
#BISCHWILLER in, 2019
GEREP$incoming[
  GEREP$Year==2019 & GEREP$name_COM == "BISCHWILLER" & GEREP$nutrient=="Matières en suspension (MES)"
  ] <- 
  GEREP$incoming[
    GEREP$Year==2019 & GEREP$name_COM == "BISCHWILLER" & GEREP$nutrient=="Matières en suspension (MES)"
    ]/10000
#BISCHWILLER out, 2019
GEREP$discharged[
  GEREP$Year==2019 & GEREP$name_COM == "BISCHWILLER" & GEREP$nutrient=="Matières en suspension (MES)"
  ] <- 
  GEREP$discharged[
    GEREP$Year==2019 & GEREP$name_COM == "BISCHWILLER" & GEREP$nutrient=="Matières en suspension (MES)"
    ]/10000
```

::: panel-tabset

### National (georisque comparison) {.tabset .tabset-fade}

The GEREP database reports higher amounts of discharged pollution in sewage networks. The explanation seems to be that the GEREP database reports more facilities than the gérosique database.

::: panel-tabset

```{r}
Source_GEREP_georisque <- "Source: géorisque and GEREP database\ncomputation by Thomas Starck"

f_graph_data_comparison <- function(
    dataset_GEREP, dataset_national_network_discharge, 
    nutri_georisq, nutri_GEREP_in, nutri_GEREP_out, nutri_nb_facilities, nutrient_label
    ){
  g1 <- ggplot(dataset_GEREP) +
    geom_area(
      data = dataset_national_network_discharge, 
      aes(Year, !!as.symbol(nutri_georisq), fill="discharged in\nsewage network"), 
      alpha=.5
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_GEREP_out), color = "discharged\nafter treatment")
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_GEREP_in), color = " discharged")
      ) +
    labs(
      x="", y="",
      color="GEREP database",
      fill="gérosique database",
      caption = Source_GEREP_georisque,
      title = "Comparison of discharge in sewage network in GEREP and géorisques databases",
      subtitle = paste0("kt", nutrient_label, " per year")
    )
  
  g2 <- ggplot(dataset_GEREP) +
    geom_area(
      data = dataset_national_network_discharge, 
      aes(Year, !!as.symbol(nutri_nb_facilities), fill="géorisque database"), 
      alpha=.5
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_nb_facilities), color = "GEREP database")
      ) +
    theme(legend.position = c(.5, .5)) +
    labs(
      x="", y="",
      color="",
      fill="",
      caption = "\n",
      title = "",
      subtitle = "Number of facilities"
    )
  
  g <- plot_grid(g1, g2, rel_widths = c(.6, .4))
  return(g)
}

Source_GEREP <- "Source: GEREP database\ncomputation by Thomas Starck"

f_graph_basin <- function(dataset, basin_select, nutri_in, nutri_out, nutri_nb_facilities, nutrient_label){
  g1 <- ggplot(dataset %>% filter(basin == basin_select)) +
    geom_line(
      aes(Year, !!as.symbol(nutri_out), color = "discharged\nafter treatment")
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_in), color = " discharged")
      ) +
    labs(
      x="", y="",
      color="",
      caption = "\n",
      title = "Pollution discharged to sewage networks",
      subtitle = paste0("kt", nutrient_label, " per year"),
    ) +
    ylim(0, NA)
  
  g2 <- ggplot(dataset %>% filter(basin == basin_select)) +
    geom_line(
      aes(Year, !!as.symbol(nutri_nb_facilities))
      ) +
    theme(legend.position = "none") +
    labs(
      x="", y="",
      title = "",
      caption = Source_GEREP,
      subtitle = "number of facilities"
    ) +
    ylim(0, NA)
  g <- plot_grid(g1, g2, rel_widths = c(.6, .4))
  return(g)
}
```

#### Pt

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "Pt", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

#### NGL

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "NGL", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

#### DBO5

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "DBO5", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

#### DCO

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "DCO", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

#### MES

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "MES", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

:::

### Basins 

::: panel-tabset

#### Pt {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "Pt_in", "Pt_out", "Pt_nb_facilities","P")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

:::

#### NGL {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

:::

#### DBO5 {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

:::

#### DCO {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

:::

#### MES {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

:::

:::

:::

## After Correction {.tabset .tabset-fade}

We recompute the aggregated flows after the correction.

::: panel-tabset

```{r}
#We compute the flows at the national scale

temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  group_by(Year, nutrient) %>%
  summarise(
    incoming = signif(sum(incoming, na.rm=T)/10^6, 3),
    discharged = signif(sum(discharged, na.rm=T)/10^6, 3),
    nb_facilities = sum(is.na(nutrient)==F)
    )

temp_in <- temp %>% select(-c(discharged, nb_facilities)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming, nb_facilities)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )

temp_nb <- temp %>% select(-c(incoming, discharged)) %>%
  spread(nutrient, nb_facilities) %>%
  rename(
    NGL_nb_facilities = `Azote total (N)`,
    Pt_nb_facilities = `Phosphore total (P)`,
    DBO5_nb_facilities = `Demande biologique en oxygène (DBO5)`,
    DCO_nb_facilities = `Demande chimique en oxygène (DCO)`,
    MES_nb_facilities = `Matières en suspension (MES)`
  )

GEREP_national <- 
  left_join(
    temp_in, temp_out, by="Year"
)
GEREP_national <-   left_join(
    GEREP_national, temp_nb, by="Year"
)

# We compute the flows for each basin
temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  group_by(Year, basin, nutrient) %>%
  summarise(
    incoming = signif(sum(incoming, na.rm=T)/10^6, 3),
    discharged = signif(sum(discharged, na.rm=T)/10^6, 3),
    nb_facilities = sum(is.na(nutrient)==F)
    )

temp_in <- temp %>% select(-c(discharged, nb_facilities)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming, nb_facilities)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )

temp_nb <- temp %>% select(-c(incoming, discharged)) %>%
  spread(nutrient, nb_facilities) %>%
  rename(
    NGL_nb_facilities = `Azote total (N)`,
    Pt_nb_facilities = `Phosphore total (P)`,
    DBO5_nb_facilities = `Demande biologique en oxygène (DBO5)`,
    DCO_nb_facilities = `Demande chimique en oxygène (DCO)`,
    MES_nb_facilities = `Matières en suspension (MES)`
  )

GEREP_basins <- left_join(
  temp_in, temp_out, by=c("Year", "basin")
)
GEREP_basins <- left_join(
  GEREP_basins, temp_nb, by=c("Year", "basin")
)

# We aslo create a the file with the flows of all the facilities
temp <- GEREP %>%
  #We only keep metropolitan France
  filter(
    basin %in% c("Adour-Garonne", "Artois-Picardie", "Loire-Bretagne", "Rhin-Meuse", "Rhône-Méditerranée", "Seine-Normandie")
    ) %>%
  mutate(ID = 1:n())

temp_in <- temp %>% select(-c(discharged)) %>%
  spread(nutrient, incoming) %>%
  rename(
    NGL_in = `Azote total (N)`,
    Pt_in = `Phosphore total (P)`,
    DBO5_in = `Demande biologique en oxygène (DBO5)`,
    DCO_in = `Demande chimique en oxygène (DCO)`,
    MES_in = `Matières en suspension (MES)`
  )

temp_out <- temp %>% select(-c(incoming)) %>%
  spread(nutrient, discharged) %>%
  rename(
    NGL_out = `Azote total (N)`,
    Pt_out = `Phosphore total (P)`,
    DBO5_out = `Demande biologique en oxygène (DBO5)`,
    DCO_out = `Demande chimique en oxygène (DCO)`,
    MES_out = `Matières en suspension (MES)`
  )


GEREP_all <- left_join(
  temp_in, 
  temp_out %>% select(ID, NGL_out, Pt_out, DBO5_out, DCO_out, MES_out), by="ID"
)
```

### National (georisque comparison) {.tabset .tabset-fade}

The GEREP database reports higher amounts of discharged pollution in sewage networks. The explanation seems to be that the GEREP database reports more facilities than the gérosique database. Contrary to the géorisque database, for now we are not able to correct the outliers in the GEREP database.

```{r}
Source_GEREP_georisque <- "Source: géorisque and GEREP database\ncomputation by Thomas Starck"

f_graph_data_comparison <- function(
    dataset_GEREP, dataset_national_network_discharge, 
    nutri_georisq, nutri_GEREP_in, nutri_GEREP_out, nutri_nb_facilities, nutrient_label
    ){
  g1 <- ggplot(dataset_GEREP) +
    geom_area(
      data = dataset_national_network_discharge, 
      aes(Year, !!as.symbol(nutri_georisq), fill="discharged in\nsewage network"), 
      alpha=.5
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_GEREP_out), color = "discharged\nafter treatment")
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_GEREP_in), color = " discharged")
      ) +
    labs(
      x="", y="",
      color="GEREP database",
      fill="gérosique database",
      caption = Source_GEREP_georisque,
      title = "Comparison of discharge in sewage network in GEREP and géorisques databases",
      subtitle = paste0("kt", nutrient_label, " per year")
    )
  
  g2 <- ggplot(dataset_GEREP) +
    geom_area(
      data = dataset_national_network_discharge, 
      aes(Year, !!as.symbol(nutri_nb_facilities), fill="géorisque database"), 
      alpha=.5
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_nb_facilities), color = "GEREP database")
      ) +
    theme(legend.position = c(.5, .5)) +
    labs(
      x="", y="",
      color="",
      fill="",
      caption = "\n",
      title = "",
      subtitle = "Number of facilities"
    )
  
  g <- plot_grid(g1, g2, rel_widths = c(.6, .4))
  return(g)
}

Source_GEREP <- "Source: GEREP database\ncomputation by Thomas Starck"

f_graph_basin <- function(dataset, basin_select, nutri_in, nutri_out, nutri_nb_facilities, nutrient_label){
  g1 <- ggplot(dataset %>% filter(basin == basin_select)) +
    geom_line(
      aes(Year, !!as.symbol(nutri_out), color = "discharged\nafter treatment")
      ) +
    geom_line(
      aes(Year, !!as.symbol(nutri_in), color = " discharged")
      ) +
    labs(
      x="", y="",
      color="",
      caption = "\n",
      title = "Pollution discharged to sewage networks",
      subtitle = paste0("kt", nutrient_label, " per year"),
    ) +
    ylim(0, NA)
  
  g2 <- ggplot(dataset %>% filter(basin == basin_select)) +
    geom_line(
      aes(Year, !!as.symbol(nutri_nb_facilities))
      ) +
    theme(legend.position = "none") +
    labs(
      x="", y="",
      title = "",
      caption = Source_GEREP,
      subtitle = "number of facilities"
    ) +
    ylim(0, NA)
  g <- plot_grid(g1, g2, rel_widths = c(.6, .4))
  return(g)
}
```

::: panel-tabset

#### Pt

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "Pt", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

#### NGL

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "NGL", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

#### DBO5

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "DBO5", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

#### DCO

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "DCO", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

#### MES

```{r}
f_graph_data_comparison(GEREP_national, national_network_discharge, "MES", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

:::

### Basins 

::: panel-tabset

#### Pt {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "Pt_in", "Pt_out", "Pt_nb_facilities","P")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "Pt_in", "Pt_out", "Pt_nb_facilities", "P")
```

:::

#### NGL {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "NGL_in", "NGL_out", "NGL_nb_facilities", "N")
```

:::

#### DBO5 {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "DBO5_in", "DBO5_out", "DBO5_nb_facilities", "DBO5")
```

:::

#### DCO {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "DCO_in", "DCO_out", "DCO_nb_facilities", "DCO")
```

:::

#### MES {.tabset .tabset-fade}

::: panel-tabset

##### Adour-Garonne

```{r}
f_graph_basin(GEREP_basins, "Adour-Garonne", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Artois-Picardie

```{r}
f_graph_basin(GEREP_basins, "Artois-Picardie", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Loire-Bretagne

```{r}
f_graph_basin(GEREP_basins, "Loire-Bretagne", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Rhin-Meuse

```{r}
f_graph_basin(GEREP_basins, "Rhin-Meuse", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Rhône-Méditerranée

```{r}
f_graph_basin(GEREP_basins, "Rhône-Méditerranée", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

##### Seine-Normandie

```{r}
f_graph_basin(GEREP_basins, "Seine-Normandie", "MES_in", "MES_out", "MES_nb_facilities", "MES")
```

:::

:::

:::

:::

# Corrections and save

Note : it is also possible to save industrial pollution at the department and regional scale.

We consider the GEREP database is more complete than the géorisque database. So we save the GEREP data.

```{r}
temp <- bind_rows(
  GEREP_national %>% mutate(basin = "Metropolitan France"),
  GEREP_basins
)

#agregated basin and national data
output_data <- "output_data/industry_network_discharge/"
f_save_csv_files(
  temp,
  output_data,
  "industry_network_discharge_GEREP_basins.csv"
)

# all facilities
f_save_csv_files(
  GEREP_all,
  output_data,
  "industry_network_discharge_GEREP_ALL.csv"
)
```

```{r remove_all_objects_in_environment}
rm(list = ls())
```
