---
title: "Sludge Composition"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F, message=F, results=F, dev='svg')


library(tidyverse) #loads multiple packages (see https://tidyverse.tidyverse.org/)

#core tidyverse packages loaded:
# ggplot2, for data visualisation. https://ggplot2.tidyverse.org/
# dplyr, for data manipulation. https://dplyr.tidyverse.org/
# tidyr, for data tidying. https://tidyr.tidyverse.org/
# readr, for data import. https://readr.tidyverse.org/
# purrr, for functional programming. https://purrr.tidyverse.org/
# tibble, for tibbles, a modern re-imagining of data frames. https://tibble.tidyverse.org/
# stringr, for strings. https://stringr.tidyverse.org/
# forcats, for factors. https://forcats.tidyverse.org/
# lubridate, for date/times. https://lubridate.tidyverse.org/

#also loads the following packages (less frequently used):
# Working with specific types of vectors:
#     hms, for times. https://hms.tidyverse.org/
# Importing other types of data:
#     feather, for sharing with Python and other languages. https://github.com/wesm/feather
#     haven, for SPSS, SAS and Stata files. https://haven.tidyverse.org/
#     httr, for web apis. https://httr.r-lib.org/
#     jsonlite for JSON. https://arxiv.org/abs/1403.2805
#     readxl, for .xls and .xlsx files. https://readxl.tidyverse.org/
#     rvest, for web scraping. https://rvest.tidyverse.org/
#     xml2, for XML. https://xml2.r-lib.org/
# Modelling
#     modelr, for modelling within a pipeline. https://modelr.tidyverse.org/
#     broom, for turning models into tidy data. https://broom.tidymodels.org/

# Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors


#setting graphs theme
theme_set(
  theme_minimal() +
    theme(
      #text = element_text(family = "Times New Roman"),
      plot.title = element_text(face="bold")
      )
  )

#setting viridis theme for colors
scale_colour_continuous <- scale_colour_viridis_c
scale_colour_discrete   <- scale_colour_viridis_d
scale_colour_binned     <- scale_colour_viridis_b
#setting viridis theme for fill
scale_fill_continuous <- scale_fill_viridis_c
scale_fill_discrete   <- scale_fill_viridis_d
scale_fill_binned     <- scale_fill_viridis_b

path_source <- "source_data/0_sludge_composition/"

Source <- "Sources: Water Agencies\nComputation Thomas Starck"

# Load the function file
source("functions.R")
```

# Esco MAFOR 

::: panel-tabset

We use a French review (collective scientific expertise) on organic fertilizers, called Esco MAFOR. [Link](https://hal.inrae.fr/hal-02796871/document) to the review (values used page 413).

## Review 

::: panel-tabset

Below are the P content, N content and N:P ratio in the reported studies, for composted and not composted urban sludge. Mohanty et al. (2006) seems to be an outlier due to very low P content and high N:P ratio. We will exclude it in the following.

```{r load_review_values}
review_sludge <- read_csv(paste0(path_source, "sludge_composition_ESCO_MAFOR.csv")) 
review_sludge$compost <- factor(
  review_sludge$compost,
  c("not composted", "composted")
)
Source <- "Source: French Collective Scientific Expertise on\nOrganic Fertilizers, 2014"
```

```{r g_composition}
g_composition <- function(dataset, select_variable){
  #order studies by increasing value
  dataset$Study <- reorder(dataset$Study, dataset %>% pull(!!as.symbol(select_variable)))
  #remove empty values
  dataset <- dataset %>% filter(is.na(!!as.symbol(select_variable))==F)
  
  ggplot(dataset) +
    geom_col(aes(Study, !!as.symbol(select_variable), fill=compost)) +
    geom_text(
      aes(Study, !!as.symbol(select_variable), label=!!as.symbol(select_variable)),
      hjust=0, family = "Times New Roman", fontface="italic"
      ) +
    coord_flip() +
    theme(
      legend.position = "none"
      ) +
    labs(
      x="", y="", 
      fill = "", 
      caption = Source
    ) +
    facet_wrap(vars(compost))
}
```

```{r rounded_values}
#we use rounded values for the graph
rounded_values <- review_sludge %>%
  mutate(
    P = round(P, 0), 
    N= round(N, 0),
    N_P_ratio = round(N_P_ratio, 1)
  )
```

### P

```{r P_content_review}
g_composition(rounded_values, "P") +
  labs(
    title = "Review of P content in urban sludge",
    subtitle = "g of P per kg of Dry Matter"
  ) +
  ylim(0, 40)
```

### N

```{r N_content_review}
g_composition(rounded_values, "N") +
  labs(
    title = "Review of N content in urban sludge",
    subtitle = "g of N per kg of Dry Matter"
  ) +
  ylim(0, 50)
```

### N:P ratio

```{r N_P_ratio_review}
g_composition(rounded_values, "N_P_ratio") +
  labs(
    title = "Review of N:P ratio in urban sludge",
    subtitle = ""
  ) +
  ylim(0, 8)
```

:::

## Mean values used for our study 

For P and N:P values, we remove the Mohanty et al. (2006) value which seems to be an outlier. Composted values are indicative, we will not use them because there are only 2 very old studies.

We compute 2 different mean values for the N:P ratio.

-   The mean of the ratios, based on the 6 values\
-   The ratio of the N and P average contents, based on 13 P values and 7 N values

For first method gices a ratio of 2.2, the second 1.5. We will use a ratio of 2 for our study.

```{r compute_mean_values}
#mean P content no composted
temp <- review_sludge %>% filter(compost=="not composted", Study!= "Mohanty et al., 2006")
mean_P_no_compost <- round(mean(temp$P, na.rm=T), 0)

#mean P content composted
temp <- review_sludge %>% filter(Study!= "Mohanty et al., 2006" & compost=="composted")
mean_P_compost <- round(mean(temp$P, na.rm=T), 0)

#mean N content not composted
temp <- review_sludge %>% filter(compost=="not composted")
mean_N_no_compost <- round(mean(temp$N, na.rm=T), 0)

#mean N content composted
temp <- review_sludge %>% filter(compost=="composted")
mean_N_compost <- round(mean(temp$N, na.rm=T), 0)

#mean N:P ratio not composted
temp <- review_sludge %>% filter(compost=="not composted", Study!= "Mohanty et al., 2006")
mean_ratio_NP_no_compost <- round(mean(temp$N_P_ratio, na.rm=T), 1)

#mean N:P ratio composted
temp <- review_sludge %>% filter(compost=="composted")
mean_ratio_NP_compost <- round(mean(temp$N_P_ratio, na.rm=T), 1)

#ratio of means N and P no compost
ratio_means_NP_no_compost <- round(mean_N_no_compost/mean_P_no_compost, 1)

#ratio of means N and P compost
ratio_means_NP_compost <- round(mean_N_compost/mean_P_compost, 1)


mean_values <- data.frame(
  compost = c("not composted", "composted"),
  P = c(mean_P_no_compost, mean_P_compost),
  N = c(mean_N_no_compost, mean_N_compost),
  N_P_mean_of_ratios = c(mean_ratio_NP_no_compost, mean_ratio_NP_compost),
  N_P_ratio_of_means = c(ratio_means_NP_no_compost, ratio_means_NP_compost),
  N_P_ratio = c(2, 0.9)
)

mean_values$compost <- factor(
  mean_values$compost,
  c("not composted", "composted")
)


f_save_csv_files(
  mean_values, 
  "output_data/sludge_composition/",
  "N_P_sludge_review.csv"
)
```

::: panel-tabset

### P

```{r}
g_composition(rounded_values %>% filter(Study!= "Mohanty et al., 2006"), "P") +
  labs(
    title = "Review of P content in urban sludge",
    subtitle = "g of P per kg of Dry Matter"
  ) +
  ylim(0, 40) +
  geom_hline(data = mean_values, aes(yintercept=P), linetype="dashed") +
  geom_text(
    data = mean_values, hjust=-0.2, family = "Times New Roman", fontface = "italic",
    aes(x="Traore et al., 1999", y=P, label=paste("mean:", P))
    )

```

### N

```{r}
g_composition(rounded_values, "N") +
  labs(
    title = "Review of N content in urban sludge",
    subtitle = "g of N per kg of Dry Matter"
  ) +
  ylim(0, 50) +
  geom_hline(data = mean_values, aes(yintercept=N), linetype="dashed") +
  geom_text(
    data = mean_values, hjust=-0.2, family = "Times New Roman", fontface = "italic",
    aes(x="Mohanty et al., 2006", y=N, label=paste("mean:", N))
    )
```

### N:P ratio

```{r}
g_composition(rounded_values %>% filter(Study!= "Mohanty et al., 2006"), "N_P_ratio") +
  labs(
    title = "Review of N:P ratio in urban sludge",
    subtitle = "we will use a N:P ratio of 2 for uncomposted sludge"
  ) +
  ylim(0, 4) +
  #mean of ratios
  geom_hline(data = mean_values, aes(yintercept=N_P_mean_of_ratios, linetype="mean of ratios")) +
  geom_text(
    data = mean_values, hjust=-0.2, family = "Times New Roman", fontface = "italic",
    aes(x=c("Sikora et al., 1982", "Traore et al., 1999"), y=N_P_mean_of_ratios, label= N_P_mean_of_ratios)
    ) +
  #ratio of means
  geom_hline(data = mean_values, aes(yintercept=N_P_ratio_of_means, linetype="ratio of means")) +
  geom_text(
    data = mean_values, hjust=1.2, family = "Times New Roman", fontface = "italic",
    aes(x=c("Sikora et al., 1982", "Traore et al., 1999"), y=N_P_ratio_of_means, label= N_P_ratio_of_means)
    ) +
  theme(legend.position = "top") +
  labs(linetype="") + 
  guides(fill = "none") #not showing composted/not composted legend
```

:::

:::

# Sillage data 

SILLAGE is a databse referencing analyses of sludge spread on crops in France. For now we cannot exploit its results, but in the future we may compare it to the method we use to determine N and P sludge content from WWTP removal efficiencies.

::: panel-tabset

## Prepare data 

::: panel-tabset

```{r}
source_data <- "source_data/0_sludge_composition/sillage_data/"
file_sillage <- read_csv2(paste0(source_data, "Export_Sillage_20230406 2.csv"))

#ATTENTION LE WEIGHT N EST PEUT ETRE PAS BON POUR LES NUTRIMENTS EN DM CAR JE DIVISE PAR DM QUI A UN N DIFFERENT DU NUTRIMENT

#infra and steu dep and name are the same, we only keep steu vector
#name
temp <- file_sillage %>%
  filter(nom_infra != nom_steu) #empty, exactly the same
#dep
temp <- file_sillage %>%
  filter(dep_infra != dep_steu) #only 18 differences, just small error

#select variables of interest and change name
sillage <- file_sillage %>%
  select(
    ID = numero_dossier,
    sludge_type = libelle_type_dossier,
    INSEE_DEP_analysis = depart_instruction,
    INSEE_DEP_WWTP = dep_steu, #same as dep_infra
    name_WWTP = nom_steu, #same as nom_infra
    code_WWTP = code_infra, 
    WWTP_or_industry = origine_infra,
    date_sample = date_prelevement,
    date_analysis = date_analyse,
    nutrient = parametre_court,
    value = resultat_analyse,
    unit = unite,
    analyse_cdn, # a comprendre
    fraction_analysee # a comprendre
  ) %>%
  mutate(
    date_analysis = as.Date(date_analysis),
    date = as.Date(date_sample),
    Year_analysis = format(date_analysis, format="%Y"),
    Year = format(date_sample, format="%Y"),
    value = as.numeric(value)
  )

#problem with years : 0017, 0018 and 1900. 
sort(unique(sillage$Year))
#we remove 1900 and transform 0017 and 0018 into 2018 and 2017
sillage <- sillage %>%
  filter(Year!="1900") %>%
  mutate(
    Year = case_when(
      Year=="0017" ~ "2017",
      Year=="0018" ~ "2018",
      T~Year
    )
  )

#add years categories
f_year_categories <- function(dataset){
  dataset <- dataset %>%
    mutate(
      Year_category = case_when(
        Year %in% c(1990, 1991, 1992, 1993, 1994) ~ "1990-1994",
        Year %in% c(1995, 1996, 1997, 1998, 1999) ~ "1995-1999",
        Year %in% c(2000, 2001, 2002, 2003, 2004) ~ "2000-2004",
        Year %in% c(2005, 2006, 2007, 2008, 2009) ~ "2005-2009",
        Year %in% c(2010, 2011, 2012, 2013, 2014) ~ "2010-2014",
        Year %in% c(2015, 2016, 2017, 2018, 2019) ~ "2015-2019",
        Year %in% c(2020, 2021, 2022, 2023, 2024) ~ "2020-2024"
      )
    )
  return(dataset)
}
sillage <- f_year_categories(sillage)

#water agency basins based on WWTP code, remove overseas
sillage <- sillage %>% 
  mutate(
    basin=
      case_when(
        substr(code_WWTP, 1, 2) == "01" ~ "Artois-Picardie",
        substr(code_WWTP, 1, 2) == "02" ~ "Rhin-Meuse",
        substr(code_WWTP, 1, 2) == "03" ~ "Seine-Normandie",
        substr(code_WWTP, 1, 2) == "04" ~ "Loire-Bretagne",
        substr(code_WWTP, 1, 2) == "05" ~ "Adour-Garonne",
        substr(code_WWTP, 1, 2) %in% c("06", "12") ~ "Rhône-Méditerranée",
        T ~ "Overseas and others"
      )
    ) %>%
  filter(basin!="Overseas and others")

sillage <- sillage %>%
  mutate(
    nutrient = case_when(
      nutrient == "K2O" ~ "K2O",
      nutrient == "NGL" ~ "NGL",
      nutrient == "Mat. sèch" ~ "DM",
      nutrient == "NK" ~ "NTK",
      nutrient == "P total" ~ "P",
      nutrient == "N orga" ~ "N_orga"
    )
  )

```

We group each WWTP by Year and take the MEDIAN value of each nutrient, and also keep in memory the number of measures. Transform K2O in K, convert from % MB to % MS

```{r}
test <- sillage %>%
  group_by(basin, code_WWTP, name_WWTP, Year_category, nutrient) %>%
  summarise(
    n = n(),
    nutrient_count = paste0("n_", nutrient),
    value = signif(median(value, na.rm = T), digits=3)
    ) %>%
  distinct() %>%
  ungroup()
test <- bind_rows(
  test %>% select(basin, code_WWTP, name_WWTP, Year_category, variable = nutrient, value),
  test %>% select(basin, code_WWTP, name_WWTP, Year_category, variable = nutrient_count, value = n)
) %>%
  spread(
    variable, value
  )
#transform K2O into K, and %MB into g per kgDM
test <- test %>% 
  rename(K = K2O, n_K = n_K2O) %>%
  mutate(K = K*0.83) %>%
  #g per kgMB
  mutate(
    K_MB = K*10, NGL_MB = NGL*10, P_MB = P*10, NTK_MB = NTK*10, N_orga_MB = N_orga/1000
  ) %>%
  #%DM
  mutate(
    K_DM = K_MB/(DM/100), NGL_DM = NGL_MB/(DM/100), P_DM = P_MB/(DM/100), NTK_DM = NTK_MB/(DM/100), N_orga_DM = N_orga_MB/(DM/100)
  )
```

first glimpse/ Only 2 WWTP for Rhin-Meuse, we remove this basin. We also remove data before 2000. carfeul N org in different unit

```{r}
f_graph_compo_all <- function(dataset, var, var_weight, exp_min, exp_max){
  g <- ggplot(dataset) +
    geom_jitter(aes(Year_category, !!as.symbol(var), size=!!as.symbol(var_weight), color=basin), alpha=.7) +
    scale_size_continuous(range = c(0, 15)) +
    scale_y_continuous(
      limits=c(10^(exp_min), 10^exp_max),
      minor_breaks = c(1:10 %o% 10^(exp_min:exp_max)),
      breaks = c(1 %o% 10^(exp_min:exp_max)),
      labels = scales::label_number(drop0trailing = TRUE)
      ) +
    coord_trans(y="log2") +
    facet_wrap(vars(basin)) +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 45, hjust=1)
      ) +
    labs(y = paste0("g", var, " per kg DM"), x="")
  return(g)
}
```

### DM

```{r}
f_graph_compo_all(test, "DM", "n_DM", -1, 2) +
  labs(y="% of MB")
```

### P 

#### g per kg MB

```{r}
f_graph_compo_all(test, "P_MB", "n_P", -1, 2)
```

#### g per kg DM

```{r}
f_graph_compo_all(test, "P_DM", "n_P", 1, 3)
```

### NGL 

#### g per kg MB

```{r}
f_graph_compo_all(test, "NGL_MB", "n_NGL", -1, 3)
```

#### g per kg DM

seems to have a problem in conversion

```{r}
f_graph_compo_all(test, "NGL_DM", "n_NGL", 0, 4)
```

### NTK 

#### g per kg MB

```{r}
f_graph_compo_all(test, "NTK_MB", "n_NTK", -1, 2)
```

#### g per kg DM

```{r}
f_graph_compo_all(test, "NTK_DM", "n_NTK", 1, 3)
```

### K 

#### g per kg MB

```{r}
f_graph_compo_all(test, "K_MB", "n_K", -2, 2)
```

#### g per kg DM

```{r}
f_graph_compo_all(test, "K_DM", "n_K", -1, 3)
```

### N orga 

#### g per kg MB

```{r}
f_graph_compo_all(test, "N_orga_MB", "n_N_orga", -1, 2)
```

#### g per kg DM

```{r}
f_graph_compo_all(test, "N_orga_DM", "n_N_orga", 1, 3)
```

### Correction

we remove Rhin-Meuse and the years before 2000

```{r}
sillage <- sillage %>% 
  filter(
    basin != "Rhin-Meuse",
    !Year_category %in% c("1990-1994", "1995-1999")
    )
test <- test %>% 
  filter(
    basin != "Rhin-Meuse",
    !Year_category %in% c("1990-1994", "1995-1999"))
```

:::

## Nutrient contents analysis 

we summarize the data, by basin and year categories, with weight nb measures for each WWTP

PROBLEM WITH NGL AND N ORGA BECAUSE FOR NOW EMPTY SUBGROUPS, PROBLEME A REGLER

```{r}
summarized <- test %>%
  group_by(basin, Year_category) %>%
  summarise(
    #P
    mean_P_MB = weighted.mean(P_MB, n_P, na.rm=T),
    med_P_MB = spatstat.geom::weighted.median(P_MB, n_P, na.rm=T),
    mean_P_DM = weighted.mean(P_DM, n_P, na.rm=T),
    med_P_DM = spatstat.geom::weighted.median(P_DM, n_P, na.rm=T),
    # #NGL
    # mean_NGL_MB = weighted.mean(NGL_MB, n_NGL, na.rm=T),
    # med_NGL_MB = spatstat::weighted.median(NGL_MB, n_NGL, na.rm=T),
    # mean_NGL_DM = weighted.mean(NGL_DM, n_NGL, na.rm=T),
    # med_NGL_DM = spatstat::weighted.median(NGL_DM, n_NGL, na.rm=T),
    #NTK
    mean_NTK_MB = weighted.mean(NTK_MB, n_NTK, na.rm=T),
    med_NTK_MB = spatstat.geom::weighted.median(NTK_MB, n_NTK, na.rm=T),
    mean_NTK_DM = weighted.mean(NTK_DM, n_NTK, na.rm=T),
    med_NTK_DM = spatstat.geom::weighted.median(NTK_DM, n_NTK, na.rm=T),
    # #N orga
    # mean_N_orga_MB = weighted.mean(N_orga_MB, n_N_orga, na.rm=T),
    # med_N_orga_MB = spatstat::weighted.median(N_orga_MB, n_N_orga, na.rm=T),
    # mean_N_orga_DM = weighted.mean(N_orga_DM, n_N_orga, na.rm=T),
    # med_N_orga_DM = spatstat::weighted.median(N_orga_DM, n_N_orga, na.rm=T),
    #K
    mean_K_MB = weighted.mean(K_MB, n_K, na.rm=T),
    med_K_MB = spatstat.geom::weighted.median(K_MB, n_K, na.rm=T),
    mean_K_DM = weighted.mean(K_DM, n_K, na.rm=T),
    med_K_DM = spatstat.geom::weighted.median(K_DM, n_K, na.rm=T),
    #DM
    mean_DM = weighted.mean(DM, n_DM, na.rm=T),
    med_DM = spatstat.geom::weighted.median(DM, n_DM, na.rm=T)
  )
```

```{r}
f_graph_compo_analysis <- function(dataset, dataset_summarised, var, var_weight, var_mean, var_med, exp_min, exp_max){
  g <- ggplot(dataset) +
    #points
    ggforce::geom_sina(
      aes(Year_category, !!as.symbol(var), size=!!as.symbol(var_weight), color=basin),
      maxwidth = .5, alpha=.7
    ) +
    #points size max
    scale_size_continuous(range = c(0, 15)) +
    #boxplot
    geom_boxplot(
      aes(Year_category, !!as.symbol(var), weight = !!as.symbol(var_weight)),
      width = 0.1,
      position = position_nudge(x = -.3, y = 0)
      ) +
    #mean
    geom_point(
      data=dataset_summarised, 
      aes(Year_category, !!as.symbol(var_mean)), 
      color="red",
      position = position_nudge(x = -.3, y = 0)
      ) +
    #uncomment to check that weigthed boxplot corresponds to weighted median
    geom_point(
      data=dataset_summarised, 
      aes(Year_category, !!as.symbol(var_med)), 
      color="red"
      ) +
    #scales
    scale_y_continuous(
      limits=c(10^(exp_min), 10^exp_max),
      minor_breaks = c(1:10 %o% 10^(exp_min:exp_max)),
      breaks = c(1 %o% 10^(exp_min:exp_max)),
      labels = scales::label_number(drop0trailing = TRUE)
      ) +
    coord_trans(y="log2") +
    facet_wrap(vars(basin)) +
    theme(
      legend.position = "none",
      axis.text.x = element_text(angle = 45, hjust=1)
      ) +
    labs(y = paste0("g", var, " per kg DM"), x="")
  
  return(g)
}
```

::: panel-tabset

### DM

```{r}
f_graph_compo_analysis(test, summarized, "DM", "n_DM", "mean_DM", "med_DM", -1, 2) +
  labs(y="% of MB")
```

### P 

#### g per kg MB

```{r}
f_graph_compo_analysis(test, summarized, "P_MB", "n_P", "mean_P_MB", "med_P_MB", -1, 2)
```

#### g per kg DM

```{r}
f_graph_compo_analysis(test, summarized, "P_DM", "n_P", "mean_P_DM", "med_P_DM", 0, 3)
```

### NGL 

#### g per kg MB

```{r}
#f_graph_compo_analysis(test, summarized, "NGL_MB", "n_NGL", "mean_NGL_MB", "med_NGL_MB", -1, 3)
```

#### g per kg DM

seems to have a problem in conversion

```{r}
#f_graph_compo_analysis(test, summarized, "NGL_DM", "n_NGL", "mean_NGL_DM", "med_NGL_DM", 0, 4)
```

### NTK 

#### g per kg MB

```{r}
f_graph_compo_analysis(test, summarized, "NTK_MB", "n_NTK", "mean_NTK_MB", "med_NTK_MB", -1, 2)
```

#### g per kg DM

```{r}
f_graph_compo_analysis(test, summarized, "NTK_DM", "n_NTK", "mean_NTK_DM", "med_NTK_DM", 0, 3)
```

### K 

#### g per kg MB

```{r}
f_graph_compo_analysis(test, summarized, "K_MB", "n_K", "mean_K_MB", "med_K_MB", -2, 2)
```

#### g per kg DM

```{r}
f_graph_compo_analysis(test, summarized, "K_DM", "n_K", "mean_K_DM", "med_K_DM", -1, 2)
```

### N orga 

#### g per kg MB

```{r}
#f_graph_compo_analysis(test, summarized, "N_orga_MB", "n_N_orga", "mean_N_orga_MB", "med_N_orga_MB", -1, 2)
```

#### g per kg DM

```{r}
#f_graph_compo_analysis(test, summarized, "N_orga_DM", "n_N_orga", "mean_N_orga_DM", "med_N_orga_DM", 1, 3)
```

:::

## Ratio

:::

```{r remove_all_objects_in_environment}
rm(list = ls())
```
